---
title: "R Analisis"
author: "Jaume Bagué Companys"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(readr)
library(dplyr)
library(biomformat)
library(readxl)
library(ggplot2)
```

# Analisis by R

## Importing from Qiime2

Our collapse table's format is in *.biom*. This fact requires a specific package for
import our database: __"biomformat"__. 
You can install it by BiocManager with: #BiocManager::install("biomformat")#

```{r}
tabla_biom="C:/Users/jaume/Documents/TFM_R/feature-table.biom" #Your path
x1 = read_biom(tabla_biom)
header(x1)
head(colnames(x1))
head(rownames(x1))
matrix_element_type(x1)
matrix_freq = as(biom_data(x1), "matrix")
```

On the other hand, list from Ancom is easier to import.

```{r}
ancom7 <- read_delim("ancom7.tsv", "\t", 
                     escape_double = FALSE, trim_ws = TRUE)
ancom7 = ancom7[1:7,1]
ancom7 = c(unlist(ancom7))
ancom7
```


## Filter table

With next code, you can filter by columns, the taxons from your ancom in your table.

```{r}
data_freq = t(matrix_freq)
prova = data_freq[,(colnames(data_freq) %in% c(ancom7))]
```

If you have problems, check:
- Any name from any object do not contain any whitespace
- Sometimes, matrix or data.frames can change **;** to **.**

This code tries to do minimal manipulation for evade this problems. 

## Join Metadata

Import metadata
```{r}
metadata_RUN456 <- read_excel("metadata_RUN456.xlsx")
```

Add a key or rule to join. In this case, it's Animal code.

```{r}
animal = c(rownames(prova))
prova = cbind.data.frame(prova, animal)
prova=left_join(prova, metadata_RUN456, by = c("animal" = "#SampleID"))
```


## Automatic Graph


Before to graph, you need to revise that your sample columns are numeric.

```{r}
str(prova)
```

If they are not numerical, you can run this loop.

```{r}
#for (i in 1:7){ # it's 1:x sample column
#prova[,i] = as.numeric(prova[,i])
#}
```

When your sample columns are numerical, you can run this loop
to graph to external pdf.


```{r}
pdf("Rplot%01d.pdf")
for (i in 1:7){
  set.seed(i)
  columnas = colnames(prova)[i]
  p = ggplot(data = prova,aes(x = Day_Status, y = prova[ ,i], col = Day_Status ))+
  geom_boxplot()+labs(title=columnas, x='Group', y='Relative Abundance' )
  f=p + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.2) +
    theme(plot.title = element_text(size = 7, face = "bold"))
  print(f)
  }
dev.off()
restoreConsole = TRUE

```


# Study focused in Day 30

In this case, Ancom from only Day 30 identified two new species
that Ancom from along time didn't do. Fractioned data from group
day 0, day 30 and day 60 will be filtered by this two species
for study their evolution.


## Importing data from Qimme2

```{r}
tabla_biom_0="C:/Users/jaume/Documents/TFM_R/feature-table_d0.biom"
tabla_biom_30="C:/Users/jaume/Documents/TFM_R/feature-table_d30.biom"
tabla_biom_60="C:/Users/jaume/Documents/TFM_R/feature-table_d60.biom"

x1 = read_biom(tabla_biom_0)
x2 = read_biom(tabla_biom_30)
x3 = read_biom(tabla_biom_60)

head(colnames(x1))
head(colnames(x2))
head(colnames(x3))

head(rownames(x1))
head(rownames(x2))
head(rownames(x3))

matrix_freq_0 = as(biom_data(x1), "matrix")
matrix_freq_30 = as(biom_data(x2), "matrix")
matrix_freq_60 = as(biom_data(x3), "matrix")

ancom_day30_focused <- read_delim("ancom_day30_focused.tsv", 
                                  "\t", escape_double = FALSE, trim_ws = TRUE)

ancom2 = ancom_day30_focused[1:2,1]
ancom2 = c(unlist(ancom2))
ancom2

```


## Filtering tables

```{r}

data_freq = t(matrix_freq_0)
prova1 = data_freq[,(colnames(data_freq) %in% c(ancom2))]
data_freq = t(matrix_freq_30)
prova2 = data_freq[,(colnames(data_freq) %in% c(ancom2))]
data_freq = t(matrix_freq_60)
prova3 = data_freq[,(colnames(data_freq) %in% c(ancom2))]

```

## Joining metadata

```{r}
animal1 = c(rownames(prova1))
animal2 = c(rownames(prova2))
animal3 = c(rownames(prova3))


prova1 = cbind.data.frame(prova1, animal1)
prova2 = cbind.data.frame(prova2, animal2)
prova3 = cbind.data.frame(prova3, animal3)



prova1=left_join(prova1, metadata_RUN456, by = c("animal1" = "#SampleID"))
prova2=left_join(prova2, metadata_RUN456, by = c("animal2" = "#SampleID"))
prova3=left_join(prova3, metadata_RUN456, by = c("animal3" = "#SampleID"))

```


## Merging tables


```{r}
prova1=prova1[,-3] ## Delete the unique different name column
prova2=prova2[,-3]
prova3=prova3[,-3]

database = rbind(prova1, prova2, prova3)

```

## Print boxplots

```{r}
pdf("Rplot%02d.pdf")
for (i in 1:2){
  set.seed(i)
  columnas = colnames(database)[i]
  p = ggplot(data = database,aes(x = Day_Status, y = database[ ,i], col = Day_Status ))+
    geom_boxplot()+labs(title=columnas, x='Group', y='Relative Abundance' )
  f=p + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.2) +
    theme(plot.title = element_text(size = 7, face = "bold"))
  print(f)
}
dev.off()
restoreConsole = TRUE
```



